---
title: "Population genetic structure of shallow and mesophotic *Montastraea cavernosa* in the Florida Keys"
author: "Alexis Sturm -- lexie.sturm@gmail.com and Ryan Eckert-- ryan.j.eckert@gmail.com"
date: "7/29/2020"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = '../data')
options(width = 88, scipen = 4)
```

***

# About this document
***  
All analyses were performed in R version 3.6.2. 
This is the code that accompanies the publication XXX. Here you will find all the code to repeat the statistical analyses performed in R and the figures created for the manuscript. The accompanying library preparation protocol and bioinformatic walkthrough can be found by clicking on the links.

If you have any issues or questions about the code please feel free to send an email to lexie.sturm@gmail.com.

***

# Basic setup of R environment
***
## Loading required packages
Make sure to set the working directory: 
`setwd("~/path/to/directory/with/data")`

```{r, set working directory}
setwd("/Users/student/Documents/GitHub/floridaKeysMcavSnp/data")
```
